Capstone Movielens Report

### 2025-05-11 reports/capstone-movielens-report.draft4.Rmd
Draft report development: function `model.tune.param_range`, `loop_starter` parameter described.

###### Sugnature
\
```{r eval=FALSE}
model.tune.param_range <- function(loop_starter,
                             tune_dir_path,
                             cache_file_base_name,
                             fn_tune.test.param_value,
                             max.identical.min_RMSE.count = 4,
                             is.cv = TRUE,
                             endpoint.min_diff = 0, #1e-07,
                             break.if_min = TRUE,
                             steps.beyond_min = 2){
  # ...
  list(best_result = param_values.best_result,
       param_values.endpoints = c(prm_val.leftmost, prm_val.rightmost, seq_increment),
       tuned.result = data.frame(parameter.value = parameter.value,
                                 RMSE = result.RMSE))
}
```

@@

  - **fn_tune.test.param_value:** A helper function that calculates the value of the `RMSE` for a given parameter value.; 
  - **break.if_min = TRUE:** A Boolean parameter that determines whether the function should terminate after completing the number of steps specified by the parameter `steps.beyond_min`, after the minimum value of the `RMSE` has been found;
  - **steps.beyond_min = 2:** (takes effect only if `break.if_min` parameter is `TRUE`) Specifies the number of steps after finding the minimum value of the `RMSE`, upon completion of which the function should terminate.

@@

###### Source Code
\
Below is the most significant part of the code of the [model.tune.param_range](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L198) function:
```{r eval=FALSE}
model.tune.param_range <- function(loop_starter,
                             tune_dir_path,
                             cache_file_base_name,
                             fn_tune.test.param_value,
                             max.identical.min_RMSE.count = 4,
                             is.cv = TRUE,
                             endpoint.min_diff = 0, #1e-07,
                             break.if_min = TRUE,
                             steps.beyond_min = 2){

@@

>>>>

###### Sugnature
\
```{r eval=FALSE}
model.tune.param_range <- function(loop_starter,
                             tune_dir_path,
                             cache_file_base_name,
                             fn_tune.test.param_value,
                             max.identical.min_RMSE.count = 4,
                             is.cv = TRUE,
                             endpoint.min_diff = 0,
                             break.if_min = TRUE,
                             steps.beyond_min = 2){
  # ...
  list(best_result = param_values.best_result,
       param_values.endpoints = c(prm_val.leftmost, prm_val.rightmost, seq_increment),
       tuned.result = data.frame(parameter.value = parameter.value,
                                 RMSE = result.RMSE))
}
```

@@

  - **loop_starter:** A numeric vector of the form `c(start, end, dvs)`, where `start` and `end` are the endpoints of the interval on which the parameter value that minimizes `RMSE` is sought.
`dvs` is a divisor for splitting the interval to transform it into a sequence of values among which the value that minimizes `RMSE` is sought.
For this purpose, the sequence step is calculated as follows:
$$
seq_step = \frac{end - start}{dvs}
$$

The sequence obtained as a result of the transformation is equivalent to the one generated by the function `seq` as follows:
```{r eval=FALSE}
seq(start, end, step)
```

  - **tune_dir_path:** ;
  - **cache_file_base_name:** ;
  - **fn_tune.test.param_value:** A helper function that calculates the value of the `RMSE` for a given parameter value.;
  - **max.identical.min_RMSE.count:** ;
  - **is.cv:** ;
  - **endpoint.min_diff:** ;
  - **break.if_min = TRUE:** A Boolean parameter that determines whether the function should terminate after completing the number of steps specified by the parameter `steps.beyond_min`, after the minimum value of the `RMSE` has been found;
  - **steps.beyond_min = 2:** (takes effect only if `break.if_min` parameter is `TRUE`) Specifies the number of steps after finding the minimum value of the `RMSE`, upon completion of which the function should terminate.

@@

###### Source Code
\
Below is the simplified version of the code of the [model.tune.param_range](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L198) function:
```{r eval=FALSE}
model.tune.param_range <- function(loop_starter,
                             tune_dir_path,
                             cache_file_base_name,
                             fn_tune.test.param_value,
                             max.identical.min_RMSE.count = 4,
                             is.cv = TRUE,
                             endpoint.min_diff = 0, #1e-07,
                             break.if_min = TRUE,
                             steps.beyond_min = 2){

@@

::: {.noteblock data-latex=""}
The full version of the code of the [model.tune.param_range](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L198) is available in the [Model Tuning](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L74) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L198) script.
:::

### 2025-05-11 reports/capstone-movielens-report.draft4.tex
Draft report development: function `model.tune.param_range`, `loop_starter` parameter described.

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-10}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model.tune.param\_range }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(loop\_starter,}
\NormalTok{                             tune\_dir\_path,}
\NormalTok{                             cache\_file\_base\_name,}
\NormalTok{                             fn\_tune.test.param\_value,}
                             \AttributeTok{max.identical.min\_RMSE.count =} \DecValTok{4}\NormalTok{,}
                             \AttributeTok{is.cv =} \ConstantTok{TRUE}\NormalTok{,}
                             \AttributeTok{endpoint.min\_diff =} \DecValTok{0}\NormalTok{, }\CommentTok{\#1e{-}07,}
                             \AttributeTok{break.if\_min =} \ConstantTok{TRUE}\NormalTok{,}
                             \AttributeTok{steps.beyond\_min =} \DecValTok{2}\NormalTok{)\{}
  \CommentTok{\# ...}
  \FunctionTok{list}\NormalTok{(}\AttributeTok{best\_result =}\NormalTok{ param\_values.best\_result,}
       \AttributeTok{param\_values.endpoints =} \FunctionTok{c}\NormalTok{(prm\_val.leftmost, prm\_val.rightmost, seq\_increment),}
       \AttributeTok{tuned.result =} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{parameter.value =}\NormalTok{ parameter.value,}
                                 \AttributeTok{RMSE =}\NormalTok{ result.RMSE))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{itemize}
\tightlist
\item
  \textbf{fn\_tune.test.param\_value:} A helper function that calculates
  the value of the \texttt{RMSE} for a given parameter value.;
\item
  \textbf{break.if\_min = TRUE:} A Boolean parameter that determines
  whether the function should terminate after completing the number of
  steps specified by the parameter \texttt{steps.beyond\_min}, after the
  minimum value of the \texttt{RMSE} has been found;
\item
  \textbf{steps.beyond\_min = 2:} (takes effect only if
  \texttt{break.if\_min} parameter is \texttt{TRUE}) Specifies the
  number of steps after finding the minimum value of the \texttt{RMSE},
  upon completion of which the function should terminate.
\end{itemize}

@@

\hfill\break
Below is the most significant part of the code of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L198}{model.tune.param\_range}
function:

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-95-1.pdf}

>>>>

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-11}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model.tune.param\_range }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(loop\_starter,}
\NormalTok{                             tune\_dir\_path,}
\NormalTok{                             cache\_file\_base\_name,}
\NormalTok{                             fn\_tune.test.param\_value,}
                             \AttributeTok{max.identical.min\_RMSE.count =} \DecValTok{4}\NormalTok{,}
                             \AttributeTok{is.cv =} \ConstantTok{TRUE}\NormalTok{,}
                             \AttributeTok{endpoint.min\_diff =} \DecValTok{0}\NormalTok{,}
                             \AttributeTok{break.if\_min =} \ConstantTok{TRUE}\NormalTok{,}
                             \AttributeTok{steps.beyond\_min =} \DecValTok{2}\NormalTok{)\{}
  \CommentTok{\# ...}
  \FunctionTok{list}\NormalTok{(}\AttributeTok{best\_result =}\NormalTok{ param\_values.best\_result,}
       \AttributeTok{param\_values.endpoints =} \FunctionTok{c}\NormalTok{(prm\_val.leftmost, prm\_val.rightmost, seq\_increment),}
       \AttributeTok{tuned.result =} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{parameter.value =}\NormalTok{ parameter.value,}
                                 \AttributeTok{RMSE =}\NormalTok{ result.RMSE))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{itemize}
\tightlist
\item
  \textbf{loop\_starter:} A numeric vector of the form
  \texttt{c(start,\ end,\ dvs)}, where \texttt{start} and \texttt{end}
  are the endpoints of the interval on which the parameter value that
  minimizes \texttt{RMSE} is sought. \texttt{dvs} is a divisor for
  splitting the interval to transform it into a sequence of values among
  which the value that minimizes \texttt{RMSE} is sought. For this
  purpose, the sequence step is calculated as follows: \[
  seq_step = \frac{end - start}{dvs}
  \]
\end{itemize}

The sequence obtained as a result of the transformation is equivalent to
the one generated by the function \texttt{seq} as follows:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{seq}\NormalTok{(start, end, step)}
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\tightlist
\item
  \textbf{tune\_dir\_path:} ;
\item
  \textbf{cache\_file\_base\_name:} ;
\item
  \textbf{fn\_tune.test.param\_value:} A helper function that calculates
  the value of the \texttt{RMSE} for a given parameter value.;
\item
  \textbf{max.identical.min\_RMSE.count:} ;
\item
  \textbf{is.cv:} ;
\item
  \textbf{endpoint.min\_diff:} ;
\item
  \textbf{break.if\_min = TRUE:} A Boolean parameter that determines
  whether the function should terminate after completing the number of
  steps specified by the parameter \texttt{steps.beyond\_min}, after the
  minimum value of the \texttt{RMSE} has been found;
\item
  \textbf{steps.beyond\_min = 2:} (takes effect only if
  \texttt{break.if\_min} parameter is \texttt{TRUE}) Specifies the
  number of steps after finding the minimum value of the \texttt{RMSE},
  upon completion of which the function should terminate.
\end{itemize}

@@

\hfill\break
Below is the simplified version of the code of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L198}{model.tune.param\_range}
function:

@@

\begin{noteblock}
The full version of the code of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L198}{model.tune.param\_range}
is available in the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L74}{Model
Tuning} section of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L198}{common-helper.functions.R}
script.

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-96-1.pdf}
